A High-Order Discontinuous Galerkin Solver with Dynamic Adaptive Mesh Refinement to Simulate Cloud Formation Processes

### 2019-09-09 writing/slides/ppam.tex
Update FV slide.
Add updatred cosine bubble.
Update stuff.

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[autostyle, english = british]{csquotes}
\usepackage[british]{babel}
\usepackage{todonotes}
\usepackage[%
  backend=biber,
  doi=false,
  url=false,
  isbn=false,
  eprint=false,
  style=verbose,
  citestyle=verbose,
  hyperref=true,
  maxnames=99,
  minnames=1,
  maxbibnames=99,
  firstinits,
  uniquename=init]{biblatex}
%\DeclareFieldFormat[inproceedings, article]{pages}{}%
%\DeclareFieldFormat[inproceedings]{organization}{}%
  % ignore some field for citations
\DeclareSourcemap{
  \maps[datatype=bibtex, overwrite]{
    \map{
      \step[fieldset=edition, null]
      \step[fieldset=publisher, null]
      \step[fieldset=pages, null]
      \step[fieldset=organization, null]
    }
  }
}

@@

  \item Constant potential temperature (temperature normalized by pressure) with larger, warm bubble and small, cold bubble on top.

  \end{itemize}
    \end{column}~%
    \begin{column}[t]{0.5\textwidth}
      \begin{figure}[h]
        \ifratio{43}{
          %\includegraphics{beamer_43_hydrostatic_equilibrium}
        } {
          %\includegraphics{beamer_169_hydrostatic_equilibrium}
        }
\caption{Background pressure in equilibrium}
\end{figure}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Two Bubbles: Simulation}
  \begin{center}
   \ifratio{43}{
%   \movie[width=0.9\textwidth, height=0.466666\textwidth, autostart,, loop, poster]{}{ppam_two_bubbles_cart.ogv}
 } {
   % todo: compute aspect ratio
   \movie[width=0.5\textwidth, height=0.31891168599\textwidth,autostart,, loop, poster]{}{ppam_two_bubbles_cart.ogv}
 }
  \end{center}
\end{frame}

@@

  \begin{equation}
 \label{eq:equation-set} 
%  \begin{array}{l}
%  \text{mass cons.} \\
%  \text{momentum cons.} \\
%  \text{energy cons.} \\
%  \text{cont.\ gas} 
% \end{array}
% :
\quad
  \pdv{}{t}
  \underbrace{
  \begin{pmatrix}
    \Qrho\\
    \Qj\\
    \QE
    \end{pmatrix}}_{\Q}
  +
  \divergence{\underbrace{
  \begin{pmatrix}
    \Qj \\
    \Qv  \otimes \Qj + \bm{I} \pressure + \textcolor{orange}{\stressT (\Q, \gradQ)}  \\
    \Qv \cdot \left(\bm{I} \QE + \bm{I} \pressure + \textcolor{orange}{\stressT (\Q, \gradQ)} \right) -
    \textcolor{orange}{\kappa \nabla T}
  \end{pmatrix}}}_{\flux(\Q, \gradQ)}
 =
  \underbrace{
  \begin{pmatrix}
    \source[\Qrho\phantom{\Qrho}]\\
    - \bm{k} \Qrho g\\
    \source[\QE]
    \end{pmatrix}}_{\source(\Q, \bm{x}, t)}
\end{equation}

@@

\textbf{Trick}: cancel out constant background pressure in flux and source

@@

\textbf{Solution}:
Modify numerical flux (Riemann solver), time step size (cfl-Condition) and boundary conditions to \textbf{allow diffusive terms}\footfullcite{dumbser2010arbitrary,gassner2008discontinuous}.

\textcolor{orange}{No explicit discretization} of gradient $\gradQ$.
\end{frame}

@@

\begin{equation}
  \label{eq:refinement-criterion}
  \operatorname{evaluate-refinement}(\Q, \mu, \sigma) =
  \begin{cases}
    \text{refine} & \text{if } \tv\left(\bm{f}(\Q)\right) \geq \mu + T_\text{refine} \sigma \\
    \text{delete} & \text{if } \tv\left(\bm{f}(\Q)\right) < \mu + T_\text{delete} \sigma \\
    \text{keep} & \text{otherwise}
    \end{cases}
\end{equation}
Choose $T_\text{refine}$ and $T_\text{delete}$ according to cost-quality trade-off.

@@

\begin{frame}
  \frametitle{AMR vs.\ Fully Refined Grid: Settings \textit{\&} Time
    To Solution}
  \begin{block}{Settings}
    \begin{description}
    \item[AMR]
      Mesh with sizes from $\SI{1000/81}{\m} = \SI{12.35}{\m}$ to $\SI{1000/3}{\m} = \SI{333.33}{\m}$.

@@

      $T_\text{refine} = 2.5$ and $T_\text{delete} = -0.5$.
    \item[Reference] $81 \times 81 = 6561$ cells.
    \end{description}
  \end{block}
  Both polynomial order 6. Viscosity $\mu = 0.01$.

  AMR grid: 1953 cells. Less than \textbf{30\%} of full grid!

  Relative $L_2$-error between AMR and fully refined: \num{2.6e-6}.
\end{frame}

\begin{frame}
  \frametitle{AMR vs.\ Fully Refined Grid: Grid}
  %  \begin{figure}[h]
  %   \centering
  %   \ifratio{43}{
  %   \movie[width=0.9\textwidth, height=0.6\textwidth, autostart,, loop, poster]{}{two_bubbles_legendre.ogv}
  % }{
  \begin{center}
   \movie[width=0.5\textwidth, height=0.318911685990\textwidth, autostart,, loop, poster]{}{ppam_two_bubbles_leg.ogv}
  \end{center}
  % }
  %   \caption{Two bubbles scenario, with \amr{}}
  % \end{figure}
\end{frame}

@@

\begin{frame}
  \frametitle{Two Bubbles: Finite Volume}
  \begin{center}
    \includegraphics[width=0.6\textwidth]{paper_two_bubbles_fv}
  \end{center}
\end{frame}

\begin{frame}{Summary}
  \begin{block}{Results}
    \begin{itemize}
    \item \textbf{ADER-DG} can be used to simulate Navier-Stokes equations.
    \item Total variation \textbf{measures edges} of numerical solution.
    \item Chebyshev's criterion finds \textbf{interesting cells}.
    \item Combination of both \textbf{accurately tracks} cloud.
    \item AMR solution is close to fully refined solution but needs \textbf{fewer} cells.
    \item \textbf{Artificial viscosity} of finite volume has a similar effect as \textbf{physical viscosity} of Navier-Stokes.  
    \end{itemize}
  % \begin{itemize}
  % \item \aderdg{} \textbf{converges} for Navier Stokes equations
  % \item \aderdg{} shows \textbf{good results} for large set of scenarios (many not shown here!)
  % \item \amr{}-criterion is \textbf{faster} and tracks cloud \textbf{correctly}.
  % \item (Limited) \aderdg{} works for reactive Euler \textit{\&} Navier-Stokes
  % \item All \textbf{extensions} to engine (diffusive fluxes, global observables) useful for \textbf{other users} of \exahype{}.
  % \end{itemize}
  \end{block}
\end{frame}

@@

\begin{frame}
  \frametitle{Shared-Memory Scaling (from\footfullcite{reinarz2019exahype})}

  \begin{center}
  \includegraphics[width=0.4\textwidth]{speedup_tbb}
  \end{center}

>>>>

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[autostyle, english = british]{csquotes}
\usepackage[british]{babel}
\usepackage{todonotes}
\usepackage[%
  backend=biber,
  doi=false,
  url=false,
  isbn=false,
  eprint=false,
  style=verbose,
  citestyle=verbose,
  hyperref=true,
  maxnames=3,
  minnames=1,
  maxbibnames=99,
  firstinits,
  uniquename=init]{biblatex}
%\DeclareFieldFormat[inproceedings, article]{pages}{}%
%\DeclareFieldFormat[inproceedings]{organization}{}%
  % ignore some field for citations
\DeclareSourcemap{
  \maps[datatype=bibtex, overwrite]{
    \map{
      \step[fieldset=edition, null]
      \step[fieldset=publisher, null]
      \step[fieldset=pages, null]
      \step[fieldset=organization, null]
    }
  }
}

@@

  \item Constant potential temperature (temperature normalized by pressure) with larger, warm bubble and small, cold bubble on top.
  \end{itemize}
    \end{column}~%
    \begin{column}[t]{0.5\textwidth}
      \begin{center}
%        \ifratio{43}{
          %\includegraphics{beamer_43_hydrostatic_equilibrium}
%        } {
          \includegraphics[trim={0 0 0 0.5cm}, clip]{ppam_hydrostatic}
%        }
Background pressure in equilibrium
\end{center}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Two Bubbles: Simulation}
  \begin{center}
   \ifratio{43}{
%   \movie[width=0.9\textwidth, height=0.466666\textwidth, autostart,, loop, poster]{}{ppam_two_bubbles_cart.ogv}
 } {
   % todo: compute aspect ratio
   %\movie[width=0.5\textwidth, height=0.31891168599\textwidth,autostart,, loop, poster]{}{ppam_two_bubbles_cart.ogv}
   \movie[width=0.6\textwidth, height=0.37571265678449256\textwidth,autostart,loop , poster]{}{ppam_two_bubbles_cart.ogv}
 }
  \end{center}
\end{frame}

@@

  \begin{equation}
 \label{eq:equation-set} 
%  \begin{array}{l}
%  \text{mass cons.} \\
%  \text{momentum cons.} \\
%  \text{energy cons.} \\
%  \text{cont.\ gas} 
% \end{array}
% :
\quad
  \pdv{}{t}
  \underbrace{
  \begin{pmatrix}
    \Qrho\\
    \Qj\\
    \QE
    \end{pmatrix}}_{\Q}
  +
  \divergence{\underbrace{
  \begin{pmatrix}
    \Qj \\
    \Qv  \otimes \Qj + \bm{I} \pressure + \textcolor{orange}{\stressT (\Q, \gradQ)}  \\
    \Qv \cdot \left(\bm{I} \QE + \bm{I} \pressure + \textcolor{orange}{\stressT (\Q, \gradQ)} \right) -
    \textcolor{orange}{\kappa \nabla T}
  \end{pmatrix}}}_{\flux(\Q, \textcolor{orange}{\gradQ})}
 =
  \underbrace{
  \begin{pmatrix}
    \source[\Qrho\phantom{\Qrho}]\\
    - \bm{k} \Qrho g\\
    \source[\QE]
    \end{pmatrix}}_{\source(\Q, \bm{x}, t)}
\end{equation}

@@

\textbf{Trick}: cancel out constant background pressure in flux and source.

@@

\textbf{Solution}:
Modify numerical flux (Riemann solver), time step size (CFL-condition) and boundary conditions to \textbf{allow diffusive terms}\footfullcite{dumbser2010arbitrary,gassner2008discontinuous}.

\textbf{No explicit discretization} of gradient \textcolor{orange}{$\gradQ$}.
\end{frame}

@@

\begin{equation}
  \label{eq:refinement-criterion}
  \operatorname{evaluate-refinement}(\Q, \mu, \sigma) =
  \begin{cases}
    \text{refine} & \text{if } \tv\left(\bm{f}(\Q)\right) \geq \mu + T_\text{refine} \sigma \\
    \text{delete} & \text{if } \tv\left(\bm{f}(\Q)\right) < \mu + T_\text{delete} \sigma \\
    \text{keep} & \text{otherwise}
    \end{cases}
\end{equation}
Choose $T_\text{refine}$ and $T_\text{delete}$ according to cost-accuracy trade-off.

@@

\begin{frame}
  \frametitle{AMR vs.\ Fully Refined Grid}
  \begin{block}{Settings}
    \begin{description}
    \item[AMR]
      Mesh with sizes from $\SI{1000/81}{\m} = \SI{12.35}{\m}$ to $\SI{1000/3}{\m} = \SI{333.33}{\m}$.

@@

      $T_\text{refine} = 2.5$ and $T_\text{delete} = -0.5$.
    \item[Reference] $81 \times 81 = 6561$ cells.
    \end{description}
  \end{block}
  Both polynomial order 6. Viscosity $\mu = 0.01$. Simulate until $t_\text{end} = \SI{600}{\s}$.

  \begin{block}{Results}
  AMR grid: 1953 cells. Less than \textbf{30\%} of full grid!

  Relative $L_2$-error between AMR and fully refined: \num{2.6e-6}.

  \end{block}
\end{frame}

\begin{frame}
  \frametitle{AMR vs.\ Fully Refined Grid: Grid}
  %  \begin{figure}[h]
  %   \centering
  %   \ifratio{43}{
  %   \movie[width=0.9\textwidth, height=0.6\textwidth, autostart,, loop, poster]{}{two_bubbles_legendre.ogv}
  % }{
  \begin{center}
   \movie[width=0.6\textwidth, height=0.382694023188\textwidth, autostart,, loop, poster]{}{ppam_two_bubbles_leg.ogv}
  \end{center}
  % }
  %   \caption{Two bubbles scenario, with \amr{}}
  % \end{figure}
\end{frame}

@@

\begin{frame}
  \frametitle{Two Bubbles: Finite Volume}
  \begin{columns}
    \begin{column}{0.6\textwidth}
  \begin{center}
    \includegraphics[width=1.0\textwidth]{paper_two_bubbles_fv}
  \end{center}
    \end{column}
    \begin{column}{0.4\textwidth}
      \begin{itemize}
      \item $7^2$ patches with $90^2$ cells each
      \item Euler equations ($\mu = 0$)
      \item Numerical viscosity smooths solution
      \end{itemize}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}{Summary}
    \begin{itemize}
    \item \textbf{ADER-DG} can be used to simulate Navier-Stokes equations.
    \item Total variation \textbf{measures edges} of numerical solution.
    \item Chebyshev's criterion finds \textbf{interesting cells}.
    \item Combination of both \textbf{accurately tracks} cloud.
    \item AMR solution is close to fully refined solution but needs \textbf{fewer} cells.
    \item \textbf{Artificial viscosity} of Finite Volume has a similar effect as \textbf{physical viscosity} of Navier-Stokes.  
    \end{itemize}
\end{frame}

@@

\begin{frame}
  \frametitle{Shared-Memory Scaling (from\footfullcite{reinarz2019exahype})}
  \begin{center}
  \includegraphics[width=0.4\textwidth]{speedup_tbb}
  \end{center}
